---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
````


## Loading and preprocessing the data
````{r}
library(tidyverse)
activity <- read_csv("activity.csv")
head(activity)
dim(activity)
str(activity)

activity$date <- as.Date(activity$date, format = "%Y-%m-%d")
```


## What is mean total number of steps taken per day?
This is the total steps taken by day (ignoring missing)

```{r}
ts <- activity |> 
        group_by(date) |> 
        summarize(total_steps = sum(steps, na.rm=TRUE))

print(ts, n=10)
```

###Histogram of the total number of steps taken each day
```{r}
ggplot(data=ts, aes(x=total_steps))+geom_histogram(bins = 61)
```
We see several days with 0 step.


## What is the average daily activity pattern?
```{r}
summary(ts$total_steps, na.rm=TRUE)
```
The average number of steps per day is 9,354
The median number of steps per day is 10,395


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
